Streaming $2$-Center with Outliers in High Dimension

### 2015-04-10 paper.tex
kiana_edition1

\section{Preliminaries}
\label{sec:pre}
Let $B(c,r)$ denote a ball with center $c$ and radius $r$. We denote by $pq$ the straight line segment between these two points and by $\card{pq}$ the length of this segment. 
%kiana: tarife delta ro chanja didam nagofte budan! benazaram ma ham mitunim nagim, baz age fek mikonin lazeme tarif konam
%Behnam: be nazaram alan bashe behtare, age doctor khast hazf mikone
Define $\delta$ of two balls $B(c,r)$ and $B'(c',r')$, the minimum distance between $B$ and $B'$ which is equal to $\max (0, \card{cc'}-r-r')$.

@@

% kiana: man be nazaram in tarif painie vase center point behtare
% Behnam: bahat moafegham

@@

\begin{algorithm}
\label{alg:1center}
\leavevmode
\begin{algorithmic}
\Function{$1$-centerWithOutliers}{$\mathcal{P}, z$}
	\State Let $p_i$ denote the elements of $\mathcal{P}$, for $i \in \set{1, \dots, n}$.
	\ForAll{$i \gets 1, \dots, z+1$}
		\State Assume $p_i$ is a non-outlier point in optimal solution.
		\State $B(c_i, r_i) \gets$ \Call{KnownPoint}{$\mathcal{P}, z, p_i$}
	\EndFor
	\State $i^* \gets \argmin_{i=1}^{z+1} r_i$
	\State \Return $B(c_{i^*}, r_{i^*})$
\EndFunction
\Statex
\Function{KnownPoint}{$\mathcal{P}, z, p$}
	\State $B \gets B(p, 0)$
	\State $Q \gets \emptyset$
	\For{$p'$ in $\mathcal{P}$}
		\If{$p' \notin B$}
			\State Insert $p'$ into $Q$
			\If{$\card{Q} = z + 1$}
				\State Remove from $Q$ a point $q$ which is closest to $p$.
				\State $B \gets B(p, \card{pq})$
			\EndIf
		\EndIf
	\EndFor
	\State \Return{B}
\EndFunction
\end{algorithmic}
\end{algorithm}
Function \textproc{KnownPoint} in Algorithm~\ref{alg:1center}, takes a point $p$ as an argument which is guaranteed to be a non-outlier point in the optimal solution. To overcome the lack of knowledge about such a point, function \textproc{$1$-centerWithOutliers} tries each of the first $z+1$ points of the stream as a candidate for a non-outlier point and generates a ball for each, using function \textproc{KnownPoint}. It returns the ball with the minimum radius. Note that, due to space limitations inherent to streaming data model, the body of for-all loop in function \textproc{$1$-centerWithOutliers} run in parallel.

@@

\begin{proof}
Let $B^*(c^*, r^*)$ be the optimal solution. Clearly, there exists a point $p$ among the first $z+1$ points of $\mathcal{P}$, which is not an outlier in the optimal solution. Since $p \in B^*$, then $\card{pp'} \le 2r^*$ for any point $p'\in B^*$. There is at least one point $q'$ among the $(z+1)$-furthest points from $p$ that is not an outlier in the optimal solution. Thus $\card{pq} \le \card{pq'} \le 2r^*$ and Algorithm~\ref{alg:1center} returns a valid solution of radius at most $2r^*$. So the proof is complete.
\end{proof}

@@

% Behnam: Aks be nazaram gooya nist ziad. va inke jaii behesh refrence dade nashode!

@@

\begin{theorem}
For any fixed $d > 1$, our single-pass data stream algorithm whose working space and update time are polynomial in $d$ and sub-linear in $n$ and returns two centers that guarantees a $(1.8 + \eps)$-approximation for the Euclidean $2$-center problem with outliers.
\end{theorem}

@@

In Section \ref{subsec:smaller}, we give a description of our algorithm for a given $r' > 0$ for the case where $\delta^* < C r^*$ and $1.2r^* \le r' < (1.2 + 2\eps/3)r^*$, our algorithm returns a $(1.8 + \eps)$-approximate solution. We explain how to find such an $r'$ and present a full description of our algorithm for the case where $\delta^* < C r^*$ in Subsection~\ref{subsec:findr}. Then we inspect the case where $\delta^* \ge C r^*$ in Subsection~\ref{subsec:bigger}.

\subsection{The Case $\delta^* < C r*$}
\label{subsec:smaller}
Our idea in this section is derived from \cite[Sec.~3.1]{ahn2014computing, kim2014improved}. To avoid duplication, we just outline the important parts of their algorithm and our modifications to it. Kim's algorithm has $10$ different states. In each step one or many of the states can be valid and their algorithm considers all of them in parallel. In each state, we have two balls. A transition between the states occurs whenever a point not covered by any of the two balls has arrived in the stream. This transition graph starts with a blank node (with no balls) and has depth four (without considering the end nodes). It is straightforward to see that this transition graph is a DAG.
Our modification is on the transition part.
Points that are covered by current solution can safely be ignored, as they do not cause any transitions in the state, and only those points that lie outside the solution are candidates for being an outlier.
In each state the number of points that should be considered as outlier is unknown. So all the possible choices should be considered. It is sufficient to have four integers, $n_1,\dots, n_4$ representing the number of outliers in depth $1$ to $4$, such that $\sum_{i=1}^{4} n_i=z$. To sum up, the modified algorithm is defined as follows:
\begin{algorithm}
\label{alg:smallc}
\leavevmode
\begin{algorithmic}
\Function{Known$r'$}{$\mathcal{P}, z, r'$}
	\For{$(n_1, \dots, n_4) \gets \set{(n_1, \dots, n_4) \provided \sum n_i=z}$}
		\State $CS \gets B(p_1,r')$.
		\State $p_{o_2} \gets $ \Call{NthOutlier}{$n_1 + 1$}
		\If{$p_{o_2}$ arrives}
			\State $CS\gets$ the sub-case of $p_{o_2} \in B_{2, 1}^*$ $\cup$ the sub-case of $p_{o_2} \in B_{2,2}^*$.
			\For{each case $S$}
				\State $p_{o_3} \gets$ \Call{NthOutlier}{$n_2 + 1$}
				\If{$p_{o_3}$ arrives} 
					\State $CS\gets CS\setminus S$ $\cup$ the sub-case of $p_{o_3} \in B_{2, 1}^*$ $\cup$ the sub-case of $p_{o_3} \in B_{2,2}^*$.
					\For {each sub-case $S'$} 
						\State $p_{o_4} \gets$ \Call{NthOutlier}{$n_3 + 1$}
						\If {$p_{o_4}$ arrives} 
							\State Replace $S'$ with a new candidate solution.
							\State $p_{o_5} \gets$ \Call{NthOutlier}{$n_4 + 1$}
							\If{$p_{o_5}$ arrives}
								\State Abandon the solution.
							\EndIf
						\EndIf
					\EndFor
				\EndIf
			\EndFor

		\EndIf
		\State $\mathop{Sol}_{n_1,n_2,n_3,n_4} \gets$ the solution with the smallest larger radius, among all the sub-cases.
	\EndFor
	\State \Return the solution $\mathop{Sol}_{n_1,n_2,n_3,n_4}$ with the smallest radius, for all valid values of $(n_1,n_2,n_3,n_4)$.
\EndFunction
\end{algorithmic}
\end{algorithm}
Here, \textproc{NthOutlier}$(n)$ returns the $n$\nth next point in $\mathcal{P}$ that does not lie in current candidate solution.

>>>>

\section{Preliminaries}
\label{sec:pre}
Let $B(c,r)$ denote a ball with center $c$ and radius $r$. We denote by $pq$ the straight line segment between these two points and by $\card{pq}$ the length of this segment. 
Define $\delta$ of two balls $B(c,r)$ and $B'(c',r')$, the minimum distance between $B$ and $B'$ which is equal to $\max (0, \card{cc'}-r-r')$.

@@

@@

\begin{algorithm}
\label{alg:1center}
\leavevmode
\begin{algorithmic}
\Function{$1$-centerWithOutliers}{$\mathcal{P}, z$}
	\State Let $p_i$ denote the elements of $\mathcal{P}$, for $i \in \set{1, \dots, n}$.
	\ForAll{$i \gets 1, \dots, z+1$}
		\State Assume $p_i$ is a non-outlier point in optimal solution.
		\State $B(c_i, r_i) \gets$ \Call{KnownPoint}{$\mathcal{P}, z, p_i$}
	\EndFor
	\State $i^* \gets \argmin_{i=1}^{z+1} r_i$
	\State \Return $B(c_{i^*}, r_{i^*})$
\EndFunction
\Statex
\Function{KnownPoint}{$\mathcal{P}, z, p$}
	\State $B \gets B(p, 0)$
	\State $Q \gets \emptyset$
	\For{$p'$ in $\mathcal{P}$}
		\If{$p' \notin B$}
			\State Insert $p'$ into $Q$
			\If{$\card{Q} = z + 1$}
				\State $q$ $\gets$ closest point to $p$ in $Q$
				\State Remove $q$ from $Q$.
				\State $B \gets B(p, \card{pq})$
			\EndIf
		\EndIf
	\EndFor
	\State \Return{B}
\EndFunction
\end{algorithmic}
\end{algorithm}
Function \textproc{KnownPoint} in Algorithm~\ref{alg:1center}, takes a point $p$ as an argument which is guaranteed to be a non-outlier point in the optimal solution. To overcome the lack of knowledge about such a point, function \textproc{$1$-centerWithOutliers} tries each of the first $z+1$ points of the stream as a candidate for a non-outlier point and generates a ball for each, using function \textproc{KnownPoint}. It returns the ball with the minimum radius. Note that, due to space limitations inherent to streaming data model, the body of for-all loop in function \textproc{$1$-centerWithOutliers} runs in parallel.

@@

\begin{proof}
Let $B^*(c^*, r^*)$ be the optimal solution. Clearly, there exists a point $p$ among the first $z+1$ points of $\mathcal{P}$, which is not an outlier in the optimal solution (Figure~\ref{fig:1center}). Since $p \in B^*$, then $\card{pp'} \le 2r^*$ for any point $p'\in B^*$. There is at least one point $q'$ among the $(z+1)$-furthest points from $p$ that is not an outlier in the optimal solution. Thus $\card{pq} \le \card{pq'} \le 2r^*$ and Algorithm~\ref{alg:1center} returns a valid solution of radius at most $2r^*$. So the proof is complete.
\end{proof}

@@

% Behnam: Aks be nazaram gooya nist ziad. va inke jaii behesh refrence dade nashode!
%kiana: hagh ba behname. akse bedune refrence nadarim asan. montaha man benazaram kheili ehtiaji nis behesh asan

@@

\begin{theorem}
For any fixed $d > 1$, our single-pass data stream algorithm whose working space and update time are polynomial in $d$ and sub-linear in $n$, returns two centers that guarantees a $(1.8 + \eps)$-approximation for the Euclidean $2$-center problem with outliers.
\end{theorem}

@@

In Section \ref{subsec:smaller}, we give a description of our algorithm for a given $r' > 0$ for the case where $\delta^* < C r^*$ and $1.2r^* \le r' < (1.2 + 2\eps/3)r^*$, which returns a $(1.8 + \eps)$-approximate solution. We explain how to find such an $r'$ and present a full description of our algorithm for the case where $\delta^* < C r^*$ in Subsection~\ref{subsec:findr}. Then we inspect the case where $\delta^* \ge C r^*$ in Subsection~\ref{subsec:bigger}.

\subsection{The Case $\delta^* < C r*$}
\label{subsec:smaller}
Our idea in this section is derived from \cite[Sec.~3.1]{ahn2014computing, kim2014improved}. To avoid duplication, we just outline the important parts of their algorithm and our modifications to it. Kim's algorithm has $10$ different states. In each step one or many of the states can be valid and their algorithm considers all of them in parallel. In each state, we have two balls. A transition between the states occurs whenever a point not covered by any of the two balls has arrived in the stream. This transition graph starts with a blank node (with no balls) and has depth four (without considering the end nodes). It is straightforward to see that this transition graph is a DAG.
Our modification is on the transition part.
Points that are covered by current solution can safely be ignored, as they do not cause any transitions in the state, and only those points that lie outside the solution are candidates for being an outlier.
In each state the number of points that should be considered as outlier is unknown. So all the possible choices should be considered. It is sufficient to have four integers, $n_1,\dots, n_4$ representing the number of outliers in depth $1$ to $4$, such that $\sum_{i=1}^{4} n_i=z$. To sum up, the modified algorithm is defined as follows:
\begin{algorithm}
\label{alg:smallc}
\leavevmode
\begin{algorithmic}
\Function{Known$r'$}{$\mathcal{P}, z, r'$}
	\For{$(n_1, \dots, n_4) \gets \set{(n_1, \dots, n_4) \provided \sum n_i=z}$}
		\State $CS \gets B(p_1,r')$.
		\State $p_{o_2} \gets $ \Call{NthOutlier}{$n_1 + 1$, $p_1$}
		\If{$p_{o_2}$ arrives}
			\State $CS\gets$ the sub-case of $p_{o_2} \in B_{2, 1}^*$ $\cup$ the sub-case of $p_{o_2} \in B_{2,2}^*$.
			\For{each case $S$}
				\State $p_{o_3} \gets$ \Call{NthOutlier}{$n_2 + 1$, $p_{o_2}$}
				\If{$p_{o_3}$ arrives} 
					\State $CS\gets CS\setminus S$ $\cup$ the sub-case of $p_{o_3} \in B_{2, 1}^*$ $\cup$ the sub-case of $p_{o_3} \in B_{2,2}^*$.
					\For {each sub-case $S'$} 
						\State $p_{o_4} \gets$ \Call{NthOutlier}{$n_3 + 1$, $p_{o_3}$}
						\If {$p_{o_4}$ arrives} 
							\State Replace $S'$ with a new candidate solution.
							\State $p_{o_5} \gets$ \Call{NthOutlier}{$n_4 + 1$, $p_{o_4}$}
							\If{$p_{o_5}$ arrives}
								\State Abandon the solution.
							\EndIf
						\EndIf
					\EndFor
				\EndIf
			\EndFor

		\EndIf
		\State $\mathop{Sol}_{n_1,n_2,n_3,n_4} \gets$ the solution with the smallest larger radius, among all the sub-cases.
	\EndFor
	\State \Return the solution $\mathop{Sol}_{n_1,n_2,n_3,n_4}$ with the smallest radius, for all valid values of $(n_1,n_2,n_3,n_4)$.
\EndFunction
\end{algorithmic}
\end{algorithm}
Here, \textproc{NthOutlier}$(n,)$ returns the $n$\nth next point in $\mathcal{P}$ that does not lie in current candidate solution.
